---
title: "Prepare file for batch import into ARW"
output: rmarkdown::html_vignette
vignette: >
  %/VignetteIndexEntry{intro}
  %/VignetteEncoding{UTF-8}
  %/VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

This vignette shows how to create the database for uploading photos in Batch Import mode into ARW. 
This vignette requires  that you have a folder that contains all the photos from a transect (or person) at a date, and *only these photos, no other photos, no subfolders.*

There are two main steps:
- Renaming the photos according to a code (optional)
- Create a data frame that will be imported into ARW
Steps where your input is needed are indicated.


# Setup
First setup the librairies. If you don't have these libraries, you can install them using `r install.packages(nameOfTheLibrary)`

```{r load necessary packages, include = FALSE}
library(tidyr)
library(dplyr)
library(magrittr)
library(here)
library(exifr)
library(stringr)
library(Photo2ARW)
```


# Photo renaming
Now, let's rename the photos. The first step is to create a table that will keep a memory of the raw photo name and the new photo name, in case we ever need to go back to the initial photos.

```{r get raw and new names}

#1 **INPUT NEEDED** 
#set the folder containing the photos to import  
pathToPhotos <- file.path(here(), "data")

#2 get the list of the photos in this folder. Sends an error if no photos are found.
(photoNames <- data.frame(rawPhotoNames = list.files(path = pathToPhotos, full.names = FALSE))) # check that these are the expected photos
nb_photos <- dim(photoNames)[1]
if (nb_photos == 0){ print("Error: no photos in this directory.")}


#3 extract the EXIF data and check that the photos contain the creation date
listEXIF <- read_exif(file.path(pathToPhotos, photoNames$rawPhotoNames))
if (!"CreateDate" %in% names(listEXIF)) { 
  print("Error: the photos do not have a CreateDate. Rename the photos manually, then use photo2batchimport to create the ARW table")
  } else if (any(is.na(listEXIF$CreateDate))){ 
    print("Error: some of the photos do not have a CreateDate. Rename the photos manually, then use photo2batchimport to create the ARW table") 
  } else { print("Photos are appropriate for automatic renaming based on CreateDate.")
    listEXIF %<>% select(FileName, CreateDate) %>% arrange(CreateDate) # sort the photos by CreateDate (chronological order)
      }
  
#4 **INPUT NEEDED** 
#determine the base code that you want to rename these photos with. The code should end with an underscore. The function will then increment a 3-digit number to each photo 
Code = "PVL_M_05_11_2024_"

#5 create a table containing the raw and new photo names. Check that everything is as expected.
(listEXIF %<>% mutate(newPhotoName = paste0(Code, str_pad(row_number(), pad = 0,width = 3 , "left"),"_ind.jpg")) %>% 
  rename(rawPhotoName = FileName)) %>% select(CreateDate, rawPhotoName, newPhotoName)

#6 **INPUT NEEDED**
#if everything is okay, store this table in a file
write.csv(listEXIF, file.path(here(), "exports", "photoNames_PVL_M_05_11_2024.csv"))
```

The second step is to actually rename the photos based on the codes that were defined.
```{r rename photos}

for (i in 1:length(photoNames$rawPhotoNames)){
  file.rename(from = file.path(pathToPhotos, photoNames$rawPhotoNames[i]), 
              to = file.path(pathToPhotos, photoNames$newPhotoNames[i]))
}
```


# Batch Import prep
```{r batch import prep}

#0 if you start here, set the path to the photos again
pathToPhotos <- file.path(here(), "data")

#1 Create a table with the photo names
photoNames <- list.files(path = pathToPhotos, full.names = FALSE)
BatchImportdf <- data.frame(Encounter.mediaAsset0 = photoNames)
print(BatchImportdf) #make sure these are the expected photos


#2 **INPUT NEEDED**
# Extract details from the photo name
# Change the submitterID to your ARW login
# if this is the first time ever that this site is prospected, and you are confident that there are no duplicates in your photos, you can assign individual names without recognition (= each photo is set to a new individual name) by setting MarkedID = T.
BatchImportdf_ready <- photo2batchimport(BatchImportdf,
                                         submitterID = "JSmith",
                                         MarkedID = F)

# Take a look at the final table
head(BatchImportdf_ready)

# Save the table for export - Change the file name to your needs
write.csv(BatchImportdf_ready, file.path(here(), "exports", "PVL_M_05112024.csv"), row.names = F)
```

Congratulations, this set of photos is ready for upload on ARW!
